---
title: "Penguins EDA"
format:
  html:
    code-fold: true
---

## Exploratory Data Analysis

```{r}
# loading libraries
library(palmerpenguins)
library(dplyr)
library(ggplot2)
library(mosaic)
library(tidyverse)
library(tidymodels)

# load in the data
df <- palmerpenguins::penguins

# Setting a theme for all the visualizations
theme_set(theme_classic())
```

```{r}
df %>%  # Start with the data frame 'df'
  group_by(species, sex) %>%  # Group data by 'species' and 'sex' columns
  summarise(
    across(
      where(is.numeric),  # Apply the following function to all numeric columns
      \(x) mean(x, na.rm = TRUE)  # Calculate the mean of each column, ignoring NA values
    )
  ) %>%
  knitr::kable()  # Convert the resulting data frame into a Markdown table using knitr::kable()

```

## Penguin Mass vs Bill Length by Species

```{r}
df %>%
  ggplot(aes(x = bill_length_mm, y = body_mass_g, color = species)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

It appears that bill length might be a determinant of body mass within a species of penguin.

## Penguin Mass vs Bill Length by Island

```{r}
df %>%
  ggplot(aes(x = bill_length_mm, y = body_mass_g, color = island)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

It appears that each island might have distinct features.

To do more analysis, let's split the data by island

```{r}
 biscoe_penguins <- penguins[penguins$island == "Biscoe", ]
dream_penguins <- penguins[penguins$island == "Dream", ]
torgersen_penguins <- penguins[penguins$island == "Torgersen", ]
```

Now, within each island, let's look at mass vs bill length by species.

## Penguin Mass vs Bill Length by Species on Biscoe Island

```{r}
biscoe_penguins %>%
  ggplot(aes(x = bill_length_mm, y = body_mass_g, color = species)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

It is interesting to note 2 things: there are only Adelie and Gentoo species penguins on Biscoe island, and it appears that bill length is a strong determinant of body mass (dependent on species.

## Penguin Mass vs Bill Length by Species on Dream Island

```{r}
dream_penguins %>%
  ggplot(aes(x = bill_length_mm, y = body_mass_g, color = species)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

While a bit spread out, there appears to be two distinct groups based on species for the penguins on dream island.

## Penguin Mass vs Bill Length by Species on Torgersen Island

```{r}
torgersen_penguins %>%
  ggplot(aes(x = bill_length_mm, y = body_mass_g, color = species)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

Only Adelie penguins live on Torgersen island.

## Penguin Mass vs Bill Length by Sex

```{r}
df %>%
  ggplot(aes(x = bill_length_mm, y = body_mass_g, color = sex)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

I was expecting bill length to be a determinant of body mass depending on sex more than it appears it might be. My hypothesis was that bill length and body mass would differ depending on sex. To test this, I will perform some additional tests:

```{r}
# Remove rows with NA values for the variables we are interested in
penguins_clean <- na.omit(df[, c("sex", "body_mass_g", "bill_length_mm")])
# Create a new variable for body mass to bill length ratio
penguins_clean <- mutate(penguins_clean, ratio = body_mass_g / bill_length_mm)

# Fit linear model
model_full <- lm(ratio ~ sex, data = penguins_clean)

# Fit a reduced model (without the sex variable)
model_reduced <- lm(ratio ~ 1, data = penguins_clean)

# Perform an F-test
test_result <- anova(model_reduced, model_full)

# Print the results of the F-test
print(test_result)
```

The null hypothesis is that the body mass to bill length ratio is not dependent on sex. The alternative hypothesis is that the body mass to bill length ratio is dependent on sex. At the .05 significance level, with a probability of nearly 0, we would reject the null hypothesis. There is evidence to indicate that sex influences the body mass to bill length ratio for penguins, regardless of their species, contrary to what the graph may indicate.

## Penguin Mass vs Bill Depth by Species

```{r}
df %>%
  ggplot(aes(x = bill_depth_mm, y = body_mass_g, color = species)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

It appears that bill depth might be a determinant of body mass depending on species. It is interesting to note that the Gentoo species appears to have a very distinct body mass to bill depth ratio than the Adelie and Chinstrap species. They appear to maybe have the same body mass to bill depth ratio.

Let's split up the data by species:

```{r}
adelie_penguins <- penguins[penguins$species == "Adelie", ]
chinstrap_penguins <- penguins[penguins$species == "Chinstrap", ]
gentoo_penguins <- penguins[penguins$species == "Gentoo", ]
```

## Adelie Penguin Mass vs Bill Length by Sex 

```{r}
adelie_penguins %>%
  ggplot(aes(x = bill_length_mm, y = body_mass_g, color = sex)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

## Chinstrap Penguin Mass vs Bill Length by Sex 

```{r}
chinstrap_penguins %>%
  ggplot(aes(x = bill_length_mm, y = body_mass_g, color = sex)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

## Gentoo Penguin Mass vs Bill Length by Sex 

```{r}
gentoo_penguins %>%
  ggplot(aes(x = bill_length_mm, y = body_mass_g, color = sex)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

```{r}
#install.packages("DBI")
#install.packages("duckdb")
```

```{python}
#import duckdb
from palmerpenguins import penguins

#con = duckdb.connect('my-db.duckdb')
#df = penguins.load_penguins()
#con.execute('CREATE TABLE penguins AS SELECT * FROM df')
#con.close()
```

```{r}
#library(DBI)
#con <- DBI::dbConnect(
#  duckdb::duckdb(), 
#  dbdir = "my-db.duckdb"
#  )
#df <- dplyr::tbl(con, "penguins")
```
